---
title: "Homework 6"
author: "PSTAT 131 John Wei"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Tree-Based Models

For this assignment, we will continue working with the file `"pokemon.csv"`, found in `/data`. The file is from Kaggle: <https://www.kaggle.com/abcsds/pokemon>.

The [Pokémon](https://www.pokemon.com/us/) franchise encompasses video games, TV shows, movies, books, and a card game. This data set was drawn from the video game series and contains statistics about 721 Pokémon, or "pocket monsters." In Pokémon games, the user plays as a trainer who collects, trades, and battles Pokémon to (a) collect all the Pokémon and (b) become the champion Pokémon trainer.

Each Pokémon has a [primary type](https://bulbapedia.bulbagarden.net/wiki/Type) (some even have secondary types). Based on their type, a Pokémon is strong against some types, and vulnerable to others. (Think rock, paper, scissors.) A Fire-type Pokémon, for example, is vulnerable to Water-type Pokémon, but strong against Grass-type.


The goal of this assignment is to build a statistical learning model that can predict the **primary type** of a Pokémon based on its generation, legendary status, and six battle statistics.

**Note: Fitting ensemble tree-based models can take a little while to run. Consider running your models outside of the .Rmd, storing the results, and loading them in your .Rmd to minimize time to knit.**

```{r}
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(klaR)
library(glmnet)
tidymodels_prefer()
Pokemon <- read_csv("Pokemon.csv")
library(janitor)
library(xgboost)
library(rpart.plot)
library(ranger)
library(vip)
library(pROC)
library("rpart.plot")
set.seed(4167)
```

### Exercise 1

Read in the data and set things up as in Homework 5:

- Use `clean_names()`
- Filter out the rarer Pokémon types
- Convert `type_1` and `legendary` to factors

Do an initial split of the data; you can choose the percentage for splitting. Stratify on the outcome variable.

Fold the training set using *v*-fold cross-validation, with `v = 5`. Stratify on the outcome variable.

Set up a recipe to predict `type_1` with `legendary`, `generation`, `sp_atk`, `attack`, `speed`, `defense`, `hp`, and `sp_def`:

- Dummy-code `legendary` and `generation`;
- Center and scale all predictors.

```{r}
pokemon <- clean_names(Pokemon)
rare_pokemon <- pokemon %>%
  filter(type_1 == 'Bug' | type_1 == 'Fire' | type_1 == 'Grass' |
  type_1 == 'Normal' | type_1 == 'Water' | type_1 == 'Psychic') %>%
  mutate(type_1 = factor(type_1), legendary = factor(legendary),
         generation = factor(generation))
pokemon_split <- initial_split(rare_pokemon, strata = type_1, prop = 0.8)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)
pokemon_fold <- vfold_cv(pokemon_train, v = 5, strata = type_1)
pokemon_recipe <- recipe(type_1 ~ legendary + generation + sp_atk
            + attack + speed + defense + hp + sp_def, data = pokemon_train) %>%
  step_dummy(c(legendary, generation)) %>%
  step_normalize(all_predictors())
```

### Exercise 2

Create a correlation matrix of the training set, using the `corrplot` package. *Note: You can choose how to handle the continuous variables for this plot; justify your decision(s).*

```{r}
library(corrplot)
corrmat_training <- pokemon_train[,sapply(pokemon_train,is.numeric)]
head(corrmat_training)
corrplot(cor(corrmat_training), method = 'number', type = 'lower') 
```

What relationships, if any, do you notice? Do these relationships make sense to you?

All the variables have positive correlation to each other to varying degrees. They also have correlation to total stats, which does make sense. Special attack, attack, and special defense have especially high correlations to total stats. 

### Exercise 3

First, set up a decision tree model and workflow. Tune the `cost_complexity` hyperparameter. Use the same levels we used in Lab 7 -- that is, `range = c(-3, -1)`. Specify that the metric we want to optimize is `roc_auc`. 

Print an `autoplot()` of the results. What do you observe? Does a single decision tree perform better with a smaller or larger complexity penalty?

```{r}
tree_spec <- decision_tree() %>%
  set_engine("rpart")
class_tree_spec <- tree_spec %>%
  set_mode("classification")
class_tree_fit <- class_tree_spec %>%
  fit(type_1 ~ legendary + generation + sp_atk +
                           attack + speed + defense + hp + sp_def,
      data = pokemon_train)
class_tree_fit %>%
  extract_fit_engine() 
```

```{r}
class_tree_wf <- workflow() %>%
  add_model(class_tree_spec %>% set_args(cost_complexity = tune())) %>%
  add_recipe(pokemon_recipe)
```

```{r}
param_grid <- grid_regular(cost_complexity(range = c(-3, -1)), levels = 10)
tune_res <- tune_grid(
  class_tree_wf, 
  resamples = pokemon_fold, 
  grid = param_grid, 
  metrics = metric_set(roc_auc)
)
```

With a single decision tree, a lower cost complexity performs better. 

### Exercise 4

What is the `roc_auc` of your best-performing pruned decision tree on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
metrics_1 <- collect_metrics(tune_res)%>%
  arrange(-mean)
```

Highest mean roc is 0.65792.

### Exercise 5

Using `rpart.plot`, fit and visualize your best-performing pruned decision tree with the *training* set.

```{r}
best_complexity <- select_best(tune_res)
class_tree_final <- finalize_workflow(class_tree_wf, best_complexity)
class_tree_final_fit <- fit(class_tree_final, data = pokemon_train)
class_tree_final_fit %>%
  extract_fit_engine() %>%
  rpart.plot()
```

### Exercise 5

Now set up a random forest model and workflow. Use the `ranger` engine and set `importance = "impurity"`. Tune `mtry`, `trees`, and `min_n`. Using the documentation for `rand_forest()`, explain in your own words what each of these hyperparameters represent.

Create a regular grid with 8 levels each. You can choose plausible ranges for each hyperparameter. Note that `mtry` should not be smaller than 1 or larger than 8. **Explain why not. What type of model would `mtry = 8` represent?**

```{r}
rf_spec <- rand_forest() %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")
rf_wf <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = tune(), trees = tune(), min_n = tune())) %>%
  add_recipe(pokemon_recipe)
```

```
pokemon_grid <- grid_regular(mtry(range = c(1, 8)), trees(range = c(1,5)), min_n(range = c(3,5)), levels = 8)
```

mtry is the number of predictors to be randomly selected in the model at each split, trees is the number of trees, and min_n is the minimal node size. 

mtry = 8 means to randomly sample 8 predictors in each split when creating the tree models.


### Exercise 6

Specify `roc_auc` as a metric. Tune the model and print an `autoplot()` of the results. What do you observe? What values of the hyperparameters seem to yield the best performance?

```{r}
tune_res_2 <- tune_grid(rf_wf, resamples = pokemon_fold, grid = pokemon_grid, metrics = metric_set(roc_auc))
autoplot(tune_res_2)
```

Node size does not seem to affect roc_auc much, but the roc_auc for the most part goes up with higher # of randomly selected predictors. Higher # of trees quite clearly create a higher roc_auc.

### Exercise 7

What is the `roc_auc` of your best-performing random forest model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
metrics <- collect_metrics(tune_res_2)
arrange(metrics, desc(mean))
```

The best performing model has roc_auc at .70616.

### Exercise 8

Create a variable importance plot, using `vip()`, with your best-performing random forest model fit on the *training* set.

Which variables were most useful? Which were least useful? Are these results what you expected, or not?

```{r}
best_complexity <- select_best(tune_res_2)
rf_final <- finalize_workflow(rf_wf, best_complexity)
rf_final_fit <- fit(rf_final, data = pokemon_train)
rf_final_fit %>%
  extract_fit_engine() %>%
  vip()
```

Special attack and attack are the highest importance, which makes sense because they also correlated with total stats earlier the most. Generations are the least important, which make sense since in a game you want different generations to be similar to be balanced.

### Exercise 9

Finally, set up a boosted tree model and workflow. Use the `xgboost` engine. Tune `trees`. Create a regular grid with 10 levels; let `trees` range from 10 to 2000. Specify `roc_auc` and again print an `autoplot()` of the results. 

What do you observe?

What is the `roc_auc` of your best-performing boosted tree model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

Finally, set up a boosted tree model and workflow. Use the `xgboost` engine. Tune `trees`. Create a regular grid with 10 levels; let `trees` range from 10 to 2000. Specify `roc_auc` and again print an `autoplot()` of the results. 

What do you observe?

What is the `roc_auc` of your best-performing boosted tree model on the folds? *Hint: Use `collect_metrics()` and `arrange()`.*

```{r}
boosted_spec <- boost_tree(trees = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
boosted_workflow <- workflow() %>%
  add_recipe(pokemon_recipe) %>%
  add_model(boosted_spec)
pokemon_grid_2 <- grid_regular(trees(range = c(10,2000)), levels = 10)
```

```{r}
tune_res_3 <- tune_grid(boosted_workflow, resamples = pokemon_fold, grid = pokemon_grid_2, metrics = metric_set(roc_auc))
autoplot(tune_res_3)
```

```{r}
metrics_3 <- collect_metrics(tune_res_3)
arrange(metrics_3, desc(mean))
```

The roc_auc of the best performing boosted tree model is .7276 at 231 trees. It peaks here and drops off with more trees.  The roc_auc flucuates at most by .0045. 

### Exercise 10

Display a table of the three ROC AUC values for your best-performing pruned tree, random forest, and boosted tree models. Which performed best on the folds? Select the best of the three and use `select_best()`, `finalize_workflow()`, and `fit()` to fit it to the *testing* set. 

Print the AUC value of your best-performing model on the testing set. Print the ROC curves. Finally, create and visualize a confusion matrix heat map.

Which classes was your model most accurate at predicting? Which was it worst at?

```{r}
prunedtree_roc <- max(metrics_1$mean)
randomforest_roc <- max(metrics$mean)
boosted_roc <- max(metrics_3$mean)
roc_values <- bind_cols(prunedtree_roc,randomforest_roc, boosted_roc)
colnames(roc_values) <-  c('prune','random forest','boosted')
roc_values
```

The best performing model is the boosted one.

```{r}
roc <- select_best(tune_res_3, metric = "roc_auc")
boosted_final <- finalize_workflow(boosted_workflow, roc)
boosted_final_fit <- fit(boosted_final, data = pokemon_test)
```

Print the AUC value of your best-performing model on the testing set. Print the ROC curves. Finally, create and visualize a confusion matrix heat map.

Which classes was your model most accurate at predicting? Which was it worst at?

```{r}
augment(boosted_final_fit, new_data = pokemon_test) %>%
  roc_curve(type_1, estimate = c(.pred_Bug, .pred_Fire,.pred_Grass, .pred_Normal, .pred_Psychic, .pred_Water)) %>%
  autoplot()
```

```{r}
augment(boosted_final_fit, new_data = pokemon_test) %>%
  conf_mat(truth = type_1, estimate = .pred_class) %>%
  autoplot(type = "heatmap")
```

The model is best at predicting normal and water and worst at fire and psychic.